---
title: "Homework 2"
output: html_document
author: "Lucía Muñoz Gil"
---

RULES:
1. Deadline for homework is Sunday 23:59. Try to turn it in at least a few minutes before to avoid any issues with failed uploads, slow internet, etc. Late admissions are not accepted.
2. Backup your work, preferrably using GitHub.
3. If you get stuck on something, try googling it first. You can find answers to a lot of questions on several public forums, but consider those to be guidelines and not solutions. ABSOLUTELY DO NOT COPY-PASTE THE ANSWERS FROM THE INTERNET.
4. If that fails, you're welcome to ask a question on the Moodle forum, but try to do it in a way which doesn't explain or hint at a part of a solution. If you cannot ask your question without revealing a part of the solution, send an e-mail to dglavas@bioinfo.hr instead of posting on the forum. 
5. ABSOLUTELY DO NOT COPY-PASTE THE ANSWERS FROM YOUR COLLEAGUES.
6. Each task states how the final solution (variable or function) should be named. Please take care to name the answers exactly as specified because it helps us in correcting your homeworks.
6. Please hand in a "clean" version of your homework. That means no error-producing code and no more than one solution for each task (unless specified otherwise in the question). Also, please leave out or at least comment parts of the code which are specific to your system (such as setting the work directory and installing packages). When you're ready to hand in the final version of your homework, please run it once more in a clean environment to make sure you have all variables defined. This means cleaning your environment (either by hand or by re-starting your R session) and then running the whole code once in the order in which it's written.
7. Do not use any additional packages besides what is in base R (so already loaded automatically when you start a new R session) unless given explicit permission to do so.


  
### 1.
##### 5 points
Create a binary operator approx (modify the name if you need to) which will compare two numeric vectors. The operator should compare numbers in the same way as ==, but with tolerance of +-0.1 (meaning TRUE if the difference in value is 0.1 or smaller, FALSE otherwise).
```{r}
#FUNCTIONS: 1º-Binary operator that compares the difference among the values in the same position of the two vectors 
`%bin_op%`<-function(vec1, vec2){
  result<-ifelse (abs(vec1 - vec2) <= 0.1, TRUE, FALSE)
  result
}

#MAIN CODE: creation of the vectors and usage of the binary operator.
vector1<-c(0, 3, 5, 6.1)
vector2<-c(0, 4, 5.05, 6)

vector1 %bin_op% vector2
```
<font color="red">You don't need ifelse() here; just " result <- abs(vec1 - vec2) <= 0.1 " would do the same thing. (This is not a mistake, just stylistic advice for code readability.) </font>



### 2.  
##### 18 points
In this task, you will continue to work on the examples from the second lecture. You can use or further modify any function we created during this lecture (any, so not just the final ones!).

A) Different coin dimensions are as follows: radii from 1.0 to 2.2 cm in increments of 0.1, thickness from 0.1 to 0.3 in increments of 0.02. Make every possible combination of those elements - so all possible pairs of radius and thickness. Save this as a single variable. Mark clearly what is radius and what is thickness.
Hint: there is a built-in R function which produces all unique combinations. It's in base R, so you shouldn't need to install any new packages. (2 points)
```{r}
#create two vectors with the different dimensions of the coin
radius<-seq(1.0, 2.2, by=0.1); radius
th<-seq(0.1, 0.3, by=0.02); th

#R function to create a dataframe with all different combinations of dimensions
dif_comb<-expand.grid("Ratio"=radius, "Thickness"=th); dif_comb
```

B) Next, create vectors of coin counts for each combination - this represents count per each type of coin per square (number of coins of one type found in one square is always between 50 and 300, chosen at random). Make the vectors differ in length because not every type of coin is found in each square. Mean vector length should be around 500, with standard deviation approximately 100. Note the "around" and "approximately" - this is not exactly normal distribution because all lengths are integers, but it should approximate normal distribution. Figure out how to do this, do not use uniform distribution!   
Save this information together with dimension combinations in one object (variable). This object can be of any type you find to be appropriate, but bear in mind that 1. you will have to work with it later and 2. it has to be clear which vector "belongs" to which radius-thickness combination. (3 points)  
```{r}
#FUNCTIONS: 1º- Function that takes a random sample of n numbers included in the interval (50-300). 2º- Forms a list with different vectors with values of first and second arguments (vectors) in the same position.

COIN_COUNT<-function(n){
  sol<-sample(c(50:300), size = n, replace = TRUE)
  sol
}

COMBINATIONS<-function(vec1, vec2){
resul<-list(c(vec1, vec2))
}


#MAIN CODE: generate a random sample of numbers that approximate to a normal distribution and that will be the different lengths of the vectors that contain coin counts for different combinations per square. For each length takes a random sample of size equal to this length of numbers (50-300) and saves it on a vector. All vectors created are saved on a list. Each radius-thickness combination is also saved on a list. Finally, the code creates a final object, which is a list, composed of the two previously mentioned lists. By doing this, each combination in list 1 is associated to a different coin count vector in list 2 by their position (from 1 to 143) in their corresponding lists.

vector_length<-round(rnorm(143, mean = 500, sd = 100))
set.seed(3)

counts<-lapply(vector_length, COIN_COUNT)


radius2<-dif_comb[, 1]
th2<-dif_comb[, 2]

comb<-mapply(COMBINATIONS, radius2, th2)

final_res<-list(comb, counts); final_res

```

<font color="red">Avoid writing functions as simple as these ones (COIN_COUNT and COMBINATIONS). It might look like a good idea at first, but if you're working on a bigger project, it starts to get really confusing, really fast, even to yourself. You can always search help if you forget what one of standard R functions does, but in order to do the same for the function you wrote, you have to find where you wrote it and decode what you meant by it at the time. Not to mention it's really hard for anyone else to follow your code. So if a function body has no more that 1 or 2 lines of code, don't declare a new function for that - just use the code in its original form.    
Good usage of lapply, but the mapply-ing COMBINATIONS is completely unnecessary. Why even split thickness and radius when you already have them in one object? Not only you could've combined all those objects (dimensions and counts' vectors) in a much more straightforward and simple way, but the end result would be much easier to handle later on as well. With the way you wrote this, you're getting a multi-listed nested list, which is a headache to work on. Even if one chose to work on a list, this one is unnecessarily complicated. I suggest you play a bit with lists and data frames, and explore in what ways you can combine different objects. It will be very useful for you later.  
Also PLEASE do not leave outputs so big to print out in the console. Use head(), View(), or something else.  
And just FYI at the end: function names are not usually in caps lock in R. Don't ask me why, but everybody does it like that and so should you (so others could read your code more easily).  
So to summarise, solution correct but code has serious problems so 1.5/3  </font>


C) Make a function which will calculate total weight of gold coins of various dimensions. The function's input should be an object of the same type and layout as the object you created in part B) - in other words, your new function should work on the object you created. The function should return one number: total weight. (4 points)
```{r}

#density of gold= 19.3 cm**3
#Being the input of the function a list, the function first extracts elements of second list (the one with the vectors for coin counts) and sums them, saving each sum in the corresponding position of the list. The list is unlisted and elements are saved in a vector Suma. For radius and thickness is similar, after second list being saved in a diferent variable, both radius and thickness of each combination are saved in different vectors. Finally, gold_coinWeight function uses every vector created in the function and calculates the total weight of all the gold coins.

gold_coinWeight <- function(n, r, th) {
    vol_one <- (r^2) * pi * th
    vol <- sum(vol_one * n)
    wgh <- (vol * 19.3) / 1000
    wgh
}

RADIUS<-function(a){
  res<-a[[1]]
}

THICK<-function(a){
  res<-a[[2]]
}
  

GOLD_COIN_WEIGHT<-function(l){
  Suma<-unlist(lapply(l[[2]], sum))
  R<-l[[1]]
  Rad<-unlist(lapply(R, RADIUS))
  Thi<-unlist(lapply(R, THICK))
  
  wgh_gold<-sum(mapply(gold_coinWeight, Suma, Rad, Thi))
  wgh_gold
}

GOLD_COIN_WEIGHT(final_res)


```

<font color="red">Same comment as above for functions RADIUS and THICKNESS. Don't write superfluous functions to avoid learning how to use apply functions properly. The listed-nested-listed monstrosity of a list you created in part B is the only only reason you need so many lapplys and functions that do basically nothing except manage said list. But you already lost points for that in B so I'm not taking them again.  
Nice usage of mapply, and GOLD_COIN_WEIGHT is a nicely written function (if you squint and ignore the list-wrangling part :)). So nice job programming but please familiarise yourself with data structures in R better!  </font>


D) Generate data for gold, silver and copper coins. Silver coins have radii from 1.5 to 3.0 in increments of 0.25 and thickness from 0.12 to 0.28 in increments of 0.02. Copper coins have radii from 1.0 to 2.2 cm in increments of 0.2 and thickness from 0.15 to 0.35 in increments of 0.05. You can assume per-square and total distributions are the same as for gold coins in B).
Basically, you need to repeat the data creation process from A) and B) three times with different parameters. Do NOT do this by writing almost the same code three times! You need to find a way to "automate" the process. You can choose to organise everything from A) and B) into one function, or split the the process into smaller parts and handle it that way - whatever you find easiest. But you need to generate all the data with one call of (each) function. You will probably use some kind of repetition or looping mechanism, but use only the ones which are optimised in R (e.g. apply family). You are not allowed to use classic for/while/repeat loops. (5 points)
```{r}

COIN_GENERATION<-function(a, b, c, d, e, f){
  radius<-seq(a, b, by=c); radius
  th<-seq(d,e, by=f); th
  vector_length<-round(rnorm(143, mean = 500, sd = 100))
  counts<-lapply(vector_length, COIN_COUNT)
  comb<-mapply(COMBINATIONS, radius, th)
  final_res<-list(comb, counts)
  final_res
}

coinWeight <- function(n, r, th, dens) {
    vol_one <- (r^2) * pi * th
    vol <- sum(vol_one * n)
    wgh <- (vol * dens) / 1000
    wgh
}

ANY_COIN_WEIGHT<-function(l){
  Suma<-unlist(lapply(l[[2]], sum))
  R<-l[[1]]
  Rad<-unlist(lapply(R, RADIUS))
  Thi<-unlist(lapply(R, THICK))
  
  wgh_gold<-sum(mapply(gold_coinWeight(dens=dens), Suma, Rad, Thi))
  wgh_gold
}


ANY_COIN_COUNTS<-function(material){
  if (material=="silver"){
    a<-1.5
    b<-3.0
    c<-0.25
    e<-0.12
    f<-0.28
    g<-0.02
    dens<-10.5
  }
  else if (material=="copper"){
    a<-1.0
    b<-2.2
    c<-0.2
    e<-0.15
    f<-0.35
    g<-0.05
    dens<-9
  }
  else if (material=="gold"){
    a<-1.0
    b<-2.2
    c<-0.1
    e<-0.1
    f<-0.3
    g<-0.02
    dens<-19.3
  }
  else break
res1<-COIN_GENERATION(a=a, b=b, c=c, e=e, f=f, g=g)
res2<-ANY_COIN_WEIGHT(res1)
res1
}

ANY_COIN_WEIGHT(silver)
```

<font color="red">This doesn't work because you hard-coded the number 143 in function COIN_GENERATION, while dimensions of silver and copper coins are such that there is a different number of combinations for them. If you corrected that, function call in the last line still wouldn't work because you treated silver as a variable and not a character string (so silver instead of "silver"). You are also missing the last, key step in this part, which is to apply ANY_COIN_WEIGHT at silver, gold and copper coin datasets simultaneously. And you're also calculating the weight but actually not exporting that but the datasets. All together, it looks to me like you ran out of time? 1.5/5 for ANY_COIN_COUNTS (althought it could be done more elegantly) and for at least trying.  </font>

E) Apply the function you created in C)* on the whole data set created in D). Calculate total weights for each type of metal separately, and the total weight you need to haul from the dragon's cave (all the coins together). You should return all that information in one object. (4 points)
*You will probably need to adjust your original function to include additional parameters. You don't need to re-write the whole function here if you don't want to, you can go back to C) and change the necessary parts there - but remember to check if it still functions correctly!
```{r}

```



### 3.  
#### 4 points
Create two functions, getSum1 and getSum2, both of which will take 2 arguments (numbers m and n) and calculate the sum of integer numbers from m to n. For example, the result of calling the function with parameters 4 and 6 should be 15 (4 + 5 + 6 = 15). You should get the same result if you called the function with 6 and 4 (in different order). Solve in two different ways.
```{r}
#FUNCTION 1: getSum1 creates a vector form 0 to the absolute value of difference between n and m, if the first number (argument) is greater than the second one is the second one which will be added to each value of the vector to get another vector with the numbers n, m and all the numbers in between to do the sum of all. If the first number is samller then it will be added to the vector.

getSum1<-function(n, m){
  l<-c(0:abs(n - m))
  res<-ifelse(n > m, sum(m + l), sum(n + l))
  res
}

getSum1(6, 4)
getSum1(4, 6)

#FUNCTION 2: getSum2 creates a sequence from first argument to second argument with a total number of points equal to the difference plus one. Then it sums every value of the vector.

getSum2<-function(n, m){
  point_number<-abs(n - m) + 1
  sequ<-seq(m, n, length.out = point_number)
  res<-sum (sequ)
  res
  }

getSum2(6, 4)
getSum2(4, 6)
```

<font color="red">nice!</font>


### 4.  
#### 5 points
You are an engineer in the Star Fleet, tasked with re-calibrating the anti-collision system. You decide to approximate spaceships as vectors in 3D coordinate space.
(You are not allowed to use function dist() in this task. Write formulas for distance calculation yourself!)

A) Create a function manhattan that calculates Manhattan distance between two vectors.
```{r}
#FUNCTION: Manhattan distance is the sum of the absolute difference between two vectors.
MANHATTAN<-function(v1, v2){
  dis<-sum(abs(v1 - v2))
  dis
}

#EXAMPLE OF USAGE
vec1<-c(0, 2, 4)
vec2<-c(-1, 2, 1)
MANHATTAN(vec1, vec2)
```

B) Create a function euclidean that calculates Euclidean distance between two vectors.
```{r}
#FUNCTION: Euclidean distance is the square root of the sum of the squared differences between two vectors
EUCLIDEAN<-function(v1, v2){
  differences<-abs(v1 - v2)**2
  dis<-sqrt(sum(differences))
  dis
}

#EXAMPLE OF USAGE
vec1<-c(0, 2, 4)
vec2<-c(-1, 2, 1)
EUCLIDEAN(vec1, vec2)
```

C) Create a function cheby that calculates Chebyshev distance between two vectors.
```{r}
#FUNCTION: Euclidean distance is the square root of the sum of the squared differences between two vectors
CHEBYSHEV<-function(v1, v2){
  difference<-abs(v1 - v2)
  dis<-max(difference)
  dis
}

#EXAMPLE OF USAGE
vec1<-c(0, 2, 4)
vec2<-c(-1, 2, 1)
CHEBYSHEV(vec1, vec2)
```

D) Create a function getDistance with three parameters (v1, v2, metric) that will calculate a distance between vectors v1 and v2. The distance to be used will be passed as a string via metric argument and can be one of the following: “Manhattan”, “Euclidean”, “Chebyshev”. Use the functions you created before inside the final function. Add a warning if distance is smaller than 3.
```{r}
#FUNCTION: depending on third argument the function calls MANHATTAN, EUCLIDEAN OR CHEBYSHEV functions in order to calculate the required distance. If the third argument is not one of this three options the code stops. Also, if the distance is smaller than 3 the code shows a warning message.

getDistance<-function(v1, v2, metric){
  if (metric == "Manhattan"){
    res<-MANHATTAN(v1, v2) 
    }
  else if (metric == "Euclidean"){
    res<-EUCLIDEAN(v1, v2)
  }
  else if (metric == "Chebyshev"){
    res<-CHEBYSHEV(v1, v2)
  }
  else break

  if (res < 3){
    warning("Distance smaller than 3!")
  }
  res
}

#EXAMPLES OF USAGE
vec1<-c(0, 2, 4)
vec2<-c(-1, 2, 1)
getDistance(vec1, vec2, "Manhattan")
getDistance(vec1, vec2, "Euclidean")
getDistance(vec1, vec2, "Chebyshev")
getDistance(vec1, vec2, "DONT KNOW")
```


### 5.  
#### 3+2 points
A) Make a function est_pi that will estimate the value of the number pi by randomly generating points inside a square and checking whether they fall within a circle inscribed in that square. Points fall within a circle if the distance from circle center is smaller than or equal to the radius of the circle.  
```{r}
#FUNCTION: est_pi 

est_pi<-function(rad){
x<-runif(100, min=-rad, max=rad)
y<-runif(100, min=-rad, max=rad)
distance<-sqrt(x**2 + y**2)

point<-ifelse(distance <= rad, "Circle", "Square")
circle_points<-as.numeric(table(point))[1]

res<-4*(circle_points / 100)
res
}

#EXAMPLES OF USAGE: 
est_pi(5)
est_pi(10)
```
<font color="red">Ok, except here we're not really intersted in circle and square dimensions (as long ast the circle is inscribed in the square). The radius doesn't matter at all since we're using the ratio of areas in our estimation. So it makes no sense at all to put radius as the input, because it doesn't influence the result at all. What does influence the precision of the estimation greatly, however, is the number of points generated. It would make much more sense to put that as the input variable.   2/3
Also, don't complicate if there's no need: sum(distance <= rad) would do the same as ifelse and table in your code, but a lot clearer and faster. </font>

B) Repeat this estimation 100,000 times using function replicate (if the computer on which you're working on can't handle it, lower the number of repetitions; if this function call runs longer than 5-10 minutes, stop it, lower the numbers, and try again). How much (in % terms) is it different from actual value of pi if you first set the seed to 1234? (Google function set.seed() and what it does.) Use the pi variable that is part of the base R package to do this comparison. 
```{r}
set.seed(1234)
new_pi<-replicate(100000, est_pi(5))
est<-(abs(mean(new_pi) - pi) / pi) * 100
est
```
<font color="red">this part is ok  2/2  </font>

<font color="red">IMPORTANT NOTICE  
Task 2 shows obvious signs of not starting your homework on time. This is a bad practice, which lost you a lot of points in this homework. Try not to do it again.  
It also bears a striking resemblance to the code one of your colleagues submitted. I don't know if you divided the tasks, worked together, or somebody copy-pasted from another, and I don't care. Next time we see something like that, everybody involved loses all points for the task, and gets additional negative points as well. This is your first and final warning. Please take it seriously.
</font>

